---
title: "Assignment1"
author: "Zach Rippas"
date: "1/26/2022"
output:
  html_document: default
  pdf_document: default
---

In this assignment, I'll be exploring the Boston census data to find something interesting.
Reading in the data set:

```{r}
library(readxl)
#options( java.parameters = "-Xmx4g" )
censusData <- read_excel("Census_Data_1880_1930.xlsx", 1)
```
More specifically, I'd look to look at ages and how age statistics have differed over time. To get the data on ages, we can take a subset of the dataset to extract only the 'year' and 'age' columns.

```{r}
ages <- subset(censusData, select = c("year", "age"))
```

Let's look at some statistics on age during the years available to us in the data. To separate the data based on age, we'll first need to know what years are recorded in the main data set.

```{r}
unique(censusData[c("year")])
```
Now we can separate the dataset into three datasets using the subset() function.

```{r}
eighteeneighty <- subset(ages, year == 1880, select = c("year", "age"))
nineteenhundred <- subset(ages, year == 1900, select = c("year", "age"))
nineteenthirty <- subset(ages, year == 1930, select = c("year", "age"))
```

Using the summary() function, we can observe some important statistics for each interval.

```{r}
# Note that we have to call the as.numeric() function on the data sets because a small amount of the data are strings (e.g. "less than 1") for age. This may slightly alter the accuracy, but not to a significant degree since these strings values are a very small subset of all the ages.

summary(as.numeric(eighteeneighty$age))
summary(as.numeric(nineteenhundred$age))
summary(as.numeric(nineteenthirty$age))
```

From this, we can observe 
